---
title: "Cyclistic portfolio project"
author: "Ihor Halas"
date: '2022-05-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Install packages and run
```{r install packages, echo=TRUE}
library(tidyverse)
library(ggplot2)  
```

## My Cyclistic portfolio project

### Step 1 Run and review our dataframe


```{r Ride data for manual and casual users }
all_2022_v2<-read_csv("all_2022_v2.csv")
head(all_2022_v2)
```

### Step 2 Lets make some calculations

#### 1.Descriptive analysis on ride_length (all figures in seconds)
```{r Descriptive analysis on ride_length}
mean(all_2022_v2$ride_length) #straight average (total ride length / rides)
median(all_2022_v2$ride_length) #midpoint number in the ascending array of ride lengths
max(all_2022_v2$ride_length) #longest ride
min(all_2022_v2$ride_length) #shortest ride
```
#### 2.Compare members and casual users
```{r Compare members and casual users}
# Compare members and casual users
aggregate(all_2022_v2$ride_length ~ all_2022_v2$member_casual, FUN = mean)
aggregate(all_2022_v2$ride_length ~ all_2022_v2$member_casual, FUN = median)
aggregate(all_2022_v2$ride_length ~ all_2022_v2$member_casual, FUN = max)
aggregate(all_2022_v2$ride_length ~ all_2022_v2$member_casual, FUN = min)
```
#### 3.See the average ride time by each day for members vs casual users
```{r average ride time by each day}
aggregate(all_2022_v2$ride_length ~ all_2022_v2$member_casual + all_2022_v2$day_of_week, FUN = mean)
```
##### 4.Analyze ridership data by type and weekday
```{r Analyze ridership data by type and weekday}
library(lubridate)
all_2022_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>%  #creates weekday field using wday()
  group_by(member_casual, weekday) %>%  #groups by user type and weekday
  summarise(number_of_rides = n()							#calculates the number of rides and average duration 
            ,average_duration = mean(ride_length)) %>% 		# calculates the average duration
  arrange(member_casual, weekday)								# sorts
```
### Step 3 Visualizations
#### Let's visualize the number of rides by rider type
```{r visualize the number of rides by rider type}
all_2022_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
```
#### Let's create a visualization for average duration
```{r visualization for average duration}
all_2022_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
```






Кінець



